{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-07-14 thesis/landslide.tex
describe sleep sets in landlside

To summarize, %for each pair of transitions $t_i$ and $t_j$ ($i<j$) in $S$,
DPOR identifies $I_{ij}$s which will (eventually) reorder each conflicting, concurrent transition pair in $S$
%executing some $t_j' t_i'$
to reach a (possibly) new program state not exposed in the current interleaving.

@@

@@

When DPOR returns no new $I_{ij}$s not already marked in the set of visited subtrees,
the exploration is complete.

>>>>

To summarize, %for each pair of transitions $t_i$ and $t_j$ ($i<j$) in $S$,
DPOR identifies $I_{ij}$s which will (eventually) reorder each conflicting, concurrent transition pair in $S$
%executing some $t_j' t_i'$
to reach a (possibly) new program state not exposed in the current interleaving.
Prior work \cite{dpor,optimal-dpor} refers to the set of these $I_{ij}$s,
for a given $i$,
as the {\em persistent set} at the preemption point after $t_{i-1}$.

@@

{\bf Further pruning.}
In the presence of non-conflicting transitions as well as conflicting ones,
DPOR's approach as described so far can still end up testing equivalent interleavings.
For example, consider the execution sequence
in Figure~\ref{fig:sleepsets}(a).
After executing this interleaving, DPOR will detect the conflict between $t_a$ and $t_c$,
and tag $I_{a2}$ (i.e., $T_2$ to run before $t_a$), causing us to next execute
Figure~\ref{fig:sleepsets}(b).
However, DPOR does not ``remember'' why it wanted to reorder $T_2$ with $T_1$
(i.e., whether $t_b$ or $t_c$ was the important one to run before $t_a$),
so from this branch,
it next tags $I_{c'1}$, causing
Figure~\ref{fig:sleepsets}(c).
to be tested next, which is regrettably equivalent to the original interleaving
($t_a$ and $t_b$ having been reordered all told, which are independent).
To recover our forgotten original purpose,
which was already fulfilled by testing the second interleaving,
DPOR can check just before tagging a new interleaving (here, $I_{c'1}$),
among all preceding transitions equivalent to the conflicting one (here, $t_{b'}$ equivalent to $t_{a'}$),
for an already-explored interleaving beginning with the target thread.
If such exists,
%then because of the depth-first search ordering,
the new interleaving is guaranteed to be equivalent to one already checked,
and can be skipped.

\begin{figure}[h]
	\begin{center}
		\begin{tabular}{p{0.3\textwidth}p{0.25\textwidth}p{0.35\textwidth}}
	\begin{tabular}{l}
	$T_1$@$t_a$: \texttt{write~x} \\
	$T_2$@$t_b$: \texttt{read~y} \\
	$T_2$@$t_c$: \texttt{read~x} \\
	\end{tabular}
	&
	\begin{tabular}{l}
	$T_2$@$t_{b'}$: \texttt{read~y} \\
	$T_2$@$t_{c'}$: \texttt{read~x} \\
	$T_1$@$t_{a'}$: \texttt{write~x} \\
	\end{tabular}
	&
	\begin{tabular}{l}
	$T_2$@$t_{b'}$: \texttt{read~y} \\
	$T_1$@$t_{a''}$: \texttt{write~x} \\
	$T_2$@$t_{c''}$: \texttt{read~x} \\
	\end{tabular}
	\\
	\\
	(a) Original interleaving.
	&
	(b) Goal interleaving.
	&
	(c) Redundant interleaving.
	\end{tabular}
	\end{center}
	\caption{Motivating example for the sleep sets optimization.
	For clarity, threads are labeled here with numbers ($T_n$) and the transitions thereby with letters ($t_i$).}
	\label{fig:sleepsets}
\end{figure}
Landslide implements this check in {\tt equiv\_already\_explored()}.
This corresponds to the {\em sleep sets} optimization described by prior work \cite{dpor,optimal-dpor},
so named because it effectively puts $T_1$ ``to sleep'' until after the true conflicting access in $t_c$.
Further optimizations such as {\em source sets} and {\em wakeup trees},
which achieve optimality (i.e., executing exactly one interleaving per equivalence class) \cite{optimal-dpor}
are not yet implemented,
although to the best of my knowledge,
they provide further reduction only in cases of 3 or more threads;
I suspect (without proof) that sleep-set DPOR is optimal for 2 threads.

% TODO: make sure it's clear that the quicksand chapter expts were run before this optmz was added,
% and that the HTM expts fro msigbovik paper were re-run for its chapater herein

@@

{\bf Termination.}
When DPOR returns no new $I_{ij}$s not already marked in the set of visited subtrees,
the exploration is complete.
